---
title: "Homework 1"
author: "Dakota Thompson"
date: "2/4/2021"
output: word_document
---

# Complete the following problems from Wickham and Grolemund:

```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Section 5.2.4, Exercise 1
### Find all flights that

```{r}
flights
```

#### Had an arrival delay of two or more hours

Arr_delay is in minutes, so 60 minutes per hour, then 2 hours would be 120 minutes. From there we have to have all of the delays that are greater than 2 hours and include the 2 hours.

```{r}
delay = filter(flights, arr_delay >= 120)
delay
```

#### Flew to Houston (IAH or HOU)

We got the destination airport of IAH or HOU.

```{r}
dest = filter(flights, dest == 'IAH' | dest == 'HOU')
dest
```

#### Were operated by United, American, or Delta

We got the carrier airline of UA (United) or AA (American) or DL (Delta).

```{r}
carrier1 = filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')
carrier1
```

#### Departed in summer (July, August, and September)

We got all the months of summer, which teh question says is July-Sept. As such we just need the montsh between 7 and 9.

```{r}
summer = filter(flights, between(month, 7, 9))
summer
```

#### Arrived more than two hours late, but didn’t leave late

We needed all the arr_delays that were more than 120 minutes late, however we also needed all the dep_delays to be less than one because we wanted 0 (left on time) and negative numbers (left early) as they didn't leave late. 

```{r}
deptimearrlate = filter(flights, arr_delay > 120 & dep_delay < 1)
deptimearrlate
```

#### Were delayed by at least an hour, but made up over 30 minutes in flight



```{r}
delaymadeback = filter(flights, dep_delay >= 60)

delaymadeback
```

#### Departed between midnight and 6am (inclusive)



```{r}
midto6 = filter(flights, )
midto6
```

## Section 5.3.1, Exercise 3
### Sort flights to find the fastest (highest speed) flights.

```{r}
flightsspeed = mutate(flights, speed = distance/air_time)
highspeedtop = arrange(flightsspeed, desc(speed))
highspeedtop
```

## Section 5.4.1, Exercise 3
### What does the any_of() function do? 

It searches a data frame for any of the characters put into the function.

### Why might it be helpful in conjunction with this vector?
#### vars <- c("year", "month", "day", "dep_delay", "arr_delay")

It could make it easier to traverse a huge data frame for the flights that aren't performing as well (being on time) and when certain flights took place. It could also be used as a way to search flights within a period of time that were late, early, or on time. 

## Section 5.5.2, Exercise 2

Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flightsarrdep = mutate(flights, arrmindep = arr_time-dep_time)
flightsarrdep = select(flightsarrdep, arrmindep, air_time)
flightsarrdep
```

I expect to see the same values as you would

## Section 5.6.7, Exercises 4 and 5

Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

## Section 5.7.1, Exercise 3

What time of day should you fly if you want to avoid delays as much as possible? 



