---
title: "Homework 1"
author: "Dakota Thompson"
date: "2/4/2021"
output: word_document
---

# Complete the following problems from Wickham and Grolemund:

```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Section 5.2.4, Exercise 1
### Find all flights that

```{r}
flights
```

#### Had an arrival delay of two or more hours

Arr_delay is in minutes, so 60 minutes per hour, then 2 hours would be 120 minutes. From there we have to have all of the delays that are greater than 2 hours and include the 2 hours.

```{r}
delay = filter(flights, arr_delay >= 120)

delay
```

#### Flew to Houston (IAH or HOU)

We got the destination airport of IAH or HOU.

```{r}
dest = filter(flights, dest == 'IAH' | dest == 'HOU')

dest
```

#### Were operated by United, American, or Delta

We got the carrier airline of UA (United) or AA (American) or DL (Delta). I also googled nycflights13 to see where to find the airline names to codes.

```{r}
airlines

carrier1 = filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')

carrier1
```

#### Departed in summer (July, August, and September)

We got all the months of summer, which the question says is July-Sept. As such we just need the months between 7 and 9.

```{r}
summer = filter(flights, between(month, 7, 9))

summer
```

#### Arrived more than two hours late, but didn’t leave late

We needed all the arr_delays that were more than 120 minutes late, however we also needed all the dep_delays to be less than one because we wanted 0 (left on time) and negative numbers (left early) as they didn't leave late. 

```{r}
deptimearrlate = filter(flights, arr_delay > 120 & dep_delay < 1)

deptimearrlate
```

#### Were delayed by at least an hour, but made up over 30 minutes in flight

We need the flights to be delayed by more than an hour in leaving, but also the ones who were delayed 30 minutes less than how they were delayed in departure to show they made that time back up.

```{r}
delaymadeback = filter(flights, arr_delay >= 120 & arr_delay <= dep_delay - 30)

delaymadeback
```

#### Departed between midnight and 6am (inclusive)

We need the times between midnight and 1 am separate from 1 to 6 because 2400 > 100 and 600. So we have to put all the times from midnight to 1 or all the times from 1 to 6. 

```{r}
midto6 = filter(flights, between(dep_time, 2400,2459) | between(dep_time, 100,600))

midto6
```

## Section 5.3.1, Exercise 3
### Sort flights to find the fastest (highest speed) flights.

I added a speed variable of distance over time. Then I arranged the data frame by speed in descending order.

```{r}
flightsspeed = mutate(flights, speed = distance/air_time)
highspeedtop = arrange(flightsspeed, desc(speed))

highspeedtop
```

## Section 5.4.1, Exercise 3
### What does the any_of() function do? 

It searches a data frame for any of the characters put into the function.

### Why might it be helpful in conjunction with this vector?
#### vars <- c("year", "month", "day", "dep_delay", "arr_delay")

It could make it easier to traverse a huge data frame for the flights that aren't performing as well (being on time) and when certain flights took place. It could also be used as a way to search flights within a period of time that were late, early, or on time. 

## Section 5.5.2, Exercise 2

Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flightsarrdep = mutate(flights, arrmindep = arr_time-dep_time)
flightsarrdep = select(flightsarrdep, arrmindep, air_time)

flightsarrdep
```

I would expect to see them to be the same because the departure and arrival times would be the beginning and end of a flight. The air time and arrival - departure are different however. To fix this I would need to account for from where they left because of time zones. 

## Section 5.6.7, Exercises 4 and 5

### Look at the number of cancelled flights per day. Is there a pattern? 

```{r}
# This takes all the flights that don't have arrivals delays or departure delays and assumes then that they were cancelled. It then groups by day to show how many were cancelled out of the total on that day.

cancelledbyday = flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(numbercancelled = sum(cancelled),
            totalflights = n())
  
cancelledbyday
```

There seems to be more cancelled flights in the winter. There even seemed to be a major snow storm in early February. 

### Is the proportion of cancelled flights related to the average delay?

```{r}
# This takes the average cancelled per day (the same way we got the total above) to compare to the mean of the delayed flights. 

cancelledvdelay = flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(
    cancelledmean = mean(cancelled), 
    departuremean = mean(dep_delay, na.rm = TRUE),
    arrivalmean = mean(arr_delay, na.rm = TRUE)) %>%
  ungroup()

cancelledvdelay
```

There seems to be higher average cancellations on days where there are higher average delays, for both departure and arrival.

### Which carrier has the worst delays? 

```{r}
# This takes all the flights by carrier to find the average arrival delay (as that matters more for next departure and could've made time back up over a flight) and then sort the average delay to have the worst on top.

worstdelay = flights %>%
  group_by(carrier) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))

worstdelay
```

It looks like F9, which is Frontier, has the worst delays.

### Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

You could compare all of the carrier delays to all the airport delays/routes. 

```{r}
# This shows how much arrival delay there is, by carrier, on each route (between airports). We took the arrival delay and the total flights for each carrier on each route, then the arrival delay just on those routes and the total amount of flights.

challengeq = flights %>%
  filter(!is.na(arr_delay)) %>%
  group_by(origin, dest, carrier) %>%
  summarise(arr_delay = sum(arr_delay),
    flights = n()) %>%
  group_by(origin, dest) %>%
  mutate(totalarrdelay = sum(arr_delay),
    totalflights = sum(flights)) %>%
  ungroup() 

challengeq

# This takes the total we made above to find the difference between the route arrival delay and the carrier arrival delay (on a route). It then takes the mean arrival delay of the carriers (arr_delay over the amount of flights) and then gets the difference between that and the difference between the route arrival delay and the carrier arrival delay (on a route). I googled how to get rid of the NaNs. Then we can isolate the actual average of the arrival delays to compare. 

challengeq2 = challengeq %>%
  mutate(otherarrdelay = (totalarrdelay - arr_delay)/(totalflights - flights),
         meanarrdelay = arr_delay / flights,
         diffarrdelay = meanarrdelay - otherarrdelay) %>%
  filter(!is.nan(diffarrdelay)) %>%
  group_by(carrier) %>%
  summarise(diffarrdelay = mean(diffarrdelay)) %>%
  arrange(desc(diffarrdelay))

challengeq2
```

## Section 5.7.1, Exercise 3

What time of day should you fly if you want to avoid delays as much as possible? 

```{r}
timeofday = flights %>%
  group_by(hour) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(arr_delay)

timeofday
```

You should fly in the early morning!